---
title: "Reproducible Research Project 1"
output: html_document
---

##  Leverage R Markdown to present Activity dataset analysis

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

## Step 1: Install needed packages and import the activity dataset

Read in the activity dataset using read.csv. Provide a quick summary of the data using str().

```{r}

data <- read.csv("activity.csv", header = TRUE, sep = ',', colClasses = c("numeric", "character","integer"))

str(data)
```


## Step 2: What is the mean / median total steps by day? Provide histogram of complete data set. 

Aggregate all steps by day then calculate the mean and median. Provide a histogram of the data for visualization.

```{r}

byday <- aggregate(steps ~ date, data, sum)
hist(byday$steps, main = paste("Steps by Day"), col="blue", xlab="Number of Steps")
actmean <- mean(byday$steps)
actmedian <- median(byday$steps)

actmean
actmedian

```

## Step 3: What is the average daily activity pattern? 

Plot mean steps (all days) by interval. Calculate max of steps by time interval. 

```{r}

interval_act <- aggregate(steps ~ interval, data, mean)

plot(interval_act$interval,interval_act$steps, type="l", xlab="Step Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")

max_interval <- interval_act[which.max(interval_act$steps),1]


```

# Step 4: Missing value analysis 

Find the total number of missing values and combine them in a plot with the valid data. Strategy for replacing missing values is the average 5-minute interval by day of week. Compare the new merged data (with NAs) mean and median and compare to the mean and median without the invalid values. Plot a histogram with both valid and NA values for visualization.

```{r}
incomplete <- sum(!complete.cases(data))
imputed_data <- transform(data, steps = ifelse(is.na(data$steps), interval_act$steps[match(data$interval, interval_act$interval)], data$steps))

imputed_data[as.character(imputed_data$date) == "2012-10-01", 1] <- 0

steps_by_day_i <- aggregate(steps ~ date, imputed_data, sum)
hist(steps_by_day_i$steps, main = paste("Total Steps Each Day"), col="red", xlab="Number of Steps")

#Create Histogram to show difference. 
hist(byday$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("red", "blue"), lwd=10)

# calculate new average and median with imputed data and compare to average and median without imputed data
mean_imp <- mean(steps_by_day_i$steps)
median_imp <- median(steps_by_day_i$steps)

mean_diff <- mean_imp - actmean
med_diff <- median_imp - actmedian

mean_diff
med_diff

```

# Step 5: Difference in activity patterns - Weekday vs. Weekend

Subset data into factors to compare weekend to weekday activity levels. Create a plot for visualization. 

```{r}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")

imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)

library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")


```