---
title: "you can change this"
author: "Zach Olivier"
date: "2/26/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output


## MY NEW SLIDE!!
wow this shit is rad

## slide with a title
#### a slide with out  a title - subheading
#### another subtittle

        - bullet list
1. ordered list
2. item b
```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```


## text options
plain text
*italicized text*
**bold text**

`for (i %in% 1:10)`

## my new slide

look at all these different options we can choose from

``` {r, comment = "$$$$", echo = T, eval = F}
head(mtcars)


```



## plots with figures
lets embed plots
```{r, echo = F, fig.align= "center", fig.cap= "check this shit out"}
plot(mtcars$wt, mtcars$mpg)



```


## if we knit this 
        -we create a html file that we can bring up into our web browser
        we can use this to push to rpubs and github
        github does alot of the formatting for you
        we can click on the html file to see the "raw" html of our html document
        we can see the actual html uusing branch called gh-pages - all that is different is a url
        it is username .github.io [path to html file]!!!
        we now can host our slides on github!!















