---
title: 'Motor Trend: Fuel Economy Study'
author: "Zach Olivier"
date: "2/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Executive Summary: Fuel Economy Study

Miles-per-Gallon (MPG) is possibly the most impactful metric used to judge new vehicles in the automotive industry. MPG is often the deciding factor in a new car purchase - higher MPG is saved money every time you fill up - driving down cost of ownership. MPG conscious consumer behavior forces automakers to produce efficient "fuel-sipping" cars. MPG is a true competitive advantage in the race for sales. 

Further, the Environmental Protection Agency (EPA) has imposed strict MPG threshold goals for all OEMs. Setting fleet-wide average MPG goals at 30+ MPG by 2022 - automakers are investing heavily in engineering to improve fuel economy of their vehicles. 

Complacency in fuel economy innovation is a death sentence in today's auto market. Customers will buy the competitor's vehicle that provides a lower lifetime cost. The EPA will fiscally punish car brands not meeting their strict MPG goals - possibly putting major players out of business. MPG is extremely important in the automotive landscape - and may even determine the fate of the business entirely. 

Today we take MPG under the microscope to examine which features of a car improve or degrade fuel economy performance. We will rigorously analyze the relationship between transmission type and MPG performance. 

        Automatic or Manual: which is better for fuel economy?

        Customer Savings: How much more will a manual transmission save you in cost of ownership?

        Industry Impact: Will producing more manual transmission goals help car brands reach the EPA's lofty goals?




### Explore: Which transmission drives MPG performance?

Surprisingly, manual transmissions tower over automatic transmission in fuel economy performance. Only two of the manual transmission vehicles sampled fell below the industry median miles per gallon.

        The mean MPG of manual transmission vehicles is 24 mpg versus just 17 for automatic. 

Intuitively - we would expect automatic transmissions to be more fuel-efficient. Drivers manually shifting gears is imperfect and these slight delays in gear shift cause more fuel consumption. Our initial findings seem to show the exact opposite. 

Is the final answer? Should you buy a manual transmission to save the most money? Should OEMs stop making automatic vehicles all together?

Not quite. Let's continue to dig deeper. 




```{r, echo = F, message=FALSE}
# install packages
library(dplyr);library(ggplot2);library(GGally);library(ggthemes)

# load the data
data(mtcars)

# format the data
mtcars <- mtcars %>% mutate(
        cyl = as.factor(cyl),
        vs = as.factor(vs),
        gear = as.factor(gear),
        carb = as.factor(carb),
        am = as.factor(ifelse(am == 0, "Automatic", "Manual"))
)

means <- aggregate(mpg ~ am, mtcars, function(i) round(mean(i)))

# plot the relationship between 
ggplot(data = mtcars) +
        geom_boxplot(aes(x = am, y = mpg, color = am, group = am)) +
        geom_point(aes(x = am, y = mpg, color = am), pch = 21, alpha = .5) +
        geom_hline(yintercept = median(mtcars$mpg), linetype = "dotted", color = "dark grey") +
        theme_tufte() +
        labs(title = "Miles-Per-Gallon Comparision: Transmission Type",
             subtitle = "Manual outperforms Automatic by a wide margin",
             color = "Transmission Type") +
        xlab("") +
        ylab("Miles-per-Gallon") +
        annotate("text", y = 20, x = 2.42, label = "Industry Median",size = 2.3,
                 color = "black") +
        theme(legend.position = "none") +
        geom_text(data = means, aes(x = am, y = mpg,label = mpg, color = am), vjust = 1)


```

### Explore: Modeling the affect of tranmission

There has to be more to the relationship between MPG and transmission. Are there other variables could help explain the difference in MPG between transmission type?

To examine this we first build a simple one variable linear regression. We regress transmission on mpg and interpret the results. 

        We see strong evidence that the effect of manual transmission is not zero (coefficient p = .000285 < .05). 
        
        The F-statistic tells us there is some overall relationship between transmission and mpg. 

        The multiple R^2 however is not strong, only ~36% of the total variance is explained by this simple model. 
        
This leads us to believe that although there is some relationship between manual transmission and mpg, we do not have the full picture. 

```{r, echo=FALSE}
# one variable model 
fit1 <- lm(mpg ~ am, data = mtcars)
summary(fit1)


```


### Explore: What variables "explain" fuel economy?

To find other variable that may better explain the relationship between MPG and Transmission we need to investigate which variables are correlated with MPG and Transmission. 

Looking down the left hand column of the pairs plot below we notice some interesting behaviors of our sample.

        Manual tranmissions in our sample have relatively low displacement values.

        Manual transmissions in our sample have relatively low horsepower values.

        Manual transmission in our sample have relatively low weight values. 

Each of these variables have high correlation with MPG - which may be biasing our results. Perhaps weight, horsepower, and displacement are the "true" indicators of mpg performance, and transmission in our simple model is getting attributed these affects. 




```{r, echo=FALSE}
# pairs plot to investigate correlation
mtcars <- mtcars %>% dplyr::select(.,-qsec, -carb, -vs, -gear)

ggpairs(mtcars, mapping = ggplot2::aes(color = am), 
        lower = list(combo=wrap("facethist", binwidth = 1)),
        title = "Miles-Per-Gallon Comparison: Transmission Type") +
        theme_tufte()



```



### Explore: Modeling the affect of multiple variables

To test our hypothesis that transmission is getting assigned the affect of other variables in our simple model - we will fit a multiple models and compare them against each other.

First we see the affect of adding displacement, weight and horsepower to the simple model. 

        Adding weight, horsepower and displacement gives us a significant gain in variance explaination (F-statistic)

Next we compare a model with all variables correlated with mpg expect transmission to a model with all variables including transmission. This will tell us if adding transmission to a model with all other variables adds significant variance explanation. 

        Adding transmission to the model does not give us a significant gain in variance explaination. (F-statistic) 



```{r, echo=FALSE}
# pairs plot to investigate correlation
fit2 <- lm(mpg ~ disp + wt + hp + am, data = mtcars)

# anova test between simple model and the "full" model
anova(fit1, fit2)

# fit without transmission variable
fit3 <- lm(mpg ~ disp + wt + hp, data = mtcars)

# comparison between "full" model and "full" model without transmission
# does adding transmission to the "full" model result in a better fit?
anova(fit3, fit2)

```



### Conclusion: Transmission Does Not Explain MPG

By including weight, horsepower and displacement in our model we see that transmission no longer gives us enough explanation of variance to keep in the model. 

        The variables of weight, horsepower and displacement are better predictors for mpg. 
        
        This matches the trend of OEMs making underpowered, lightweight vehicles to achieve better MPG performance. 

Didn't my simple model show there was some affect of transmission on mpg? Yes. 

However, the relationship did not explain enough variance after the other highly correlated variables were included in the model. The F statistic from our anova results is based on the number of predictors added to the model - the added variable needs to overcome a certain threshold of variance explained to be included over the penalty of adding another predictor. In our case, transmission did not overcome this threshold. 

We conclude by adding that automakers should focus on weight, horsepower and cylinders to meet their internal and external MPG goals. 

        Each unit increase in weight reduces mpg performance by ~4 mpg holding displacement and horsepower constant.  

```{r, echo=FALSE}
# fit without transmission variable
fit3 <- lm(mpg ~ disp + wt + hp, data = mtcars)
summary(fit3)$coef

# model diagnostics
par(mfrow = c(2,2))
plot(fit3)

```





















